### 2014-09-06 relax2.tex
relax2 typo fixed

In particular, we have the following criteria in determining if an 
unexecuted branch should be included in the constraint system. These are 
essentially the bound of our path relaxation.
(1) If there is a loop inside the unexecuted branch, the loop bound needs to
be resolved to a constant from the trace (otherwise we do not know how many
times we need to unroll). (2) The branch
is ignored if it contains recursive function(s). (3) Every base reference 
variable in the branch needs to be fully resolved to a concrete object,
as required by our earlier analysis (Section~\ref{sec:relax1}). This can be 
done by backward traversal of the trace:
if the base reference depends on some shared read in the explored 
branch, of which the value is uncertain, it cannot be 
resolved to an object; if the base reference depends on multiple definitions 
from different paths, it cannot be resolved to an object.
Otherwise, it can be resolved.
(4) The indices of array read accesses in the branch 
need to be resolved to constant values from the trace (with the preceding unexecuted branch extensions). This is to
prevent reading an array element that was not defined. 
(5) The branch is ignored if it uses collections. 
(6) For method calls, we require the  caller object can be fully resolved from
the trace.

%We start with a branch event $e_b$ in the trace, without loss of generality, 
%suppose its true branch is taken, we need to explore the false branch 
%(thread-locally with the same thread as $e_b$) and add constraints to model it.
% As the static analysis has some known challenges, we make several 
%simplifying assumptions: 
%(1) the unexplored branch should not contain 
%loops or recursions because the loop bound may depend on some shared variable 
%which is statically unknown; (2) every base reference variable should be 
%fully resolved to a concrete object, as required by our analysis 
%(Section~\ref{sec:relax1}). We apply dataflow analysis (reaching definitions) 
%to judge: If the base reference depends on some shared read in the explored 
%branch, of which the value is statically unknown, it cannot be statically 
%resolved to an object; If the base reference depends on multiple definitions 
%from different paths, it cannot be statically resolved to an object. 
%Otherwise, it can be resolved. (3) We assume the unexplored branch does not 
%contain array access because the array accesses often use the index 
%expressions that are statically unknown. Similarly, as the collections 
%often use array for implementation, we assume they are also absent. (4) We 
%inline the method calls (for two levels) in the unexplored branch, similar 
%to the base object, we assume the caller object can be fully resolved 
%statically.  If any of the above assumptions is violated, we  avoid exploring 
%the unexplored branches and adopt the purely dynamic solution in Section~\ref{sec:relax1}. 
Next, we will explain how to encode an unexecuted branch in details, 
using the example in Figure~\ref{fig:running} for demonstration.
We assume one of the branches of event $e_b$ needs to be extended in
the following discussion.

@@

{\bf SSA Form\ }  Similar to in Section~\ref{sec:relax1}, we need the SSA 
form of the instruction in each event. To achieve this, we apply SSA 
transformation to the code of unexecuted branch and combine its SSA form 
with the existing SSA form of the corresponding executed branch to produce the final 
SSA form. The SSA transformation is standard, but special considerations
are needed.
%issues need to be considered.
\begin{itemize}
\item {\bf Local primitive variables\ }  
For local reads inside the extended branch whose definitions are
%Some local variable is defined 
before the branch event $e_b$, we reuse their SSA forms which 
%of which the SSA form 
are already in the trace $\tau$.
%, we reuse the SSA form.  
In addition, 
%the SSA transformation 
we encode conditional statements inside the extended branch using
Phi node, e.g., $y=Phi(y_1, C_1, y_2, C_2, \dots y_i, C_i)$, 
meaning that $y$ is equal to $y_i$ in the condition $C_i$.
Note that we do not require paths inside the extended branch be fully
resolved. 
Phi nodes are also used to integrate the definitions in the extended
branch with the executed branch.

@@

{\bf Intra-thread Value Constraint \ } Intra-thread value constraint is the 
core constraint that realizes the exploration of different branches. It 
asserts the value constraints imposed by the {\sf assign/heapr/heapw} events 
take effect only when the guarding condition is satisfied. 
Suppose $IF(e_b)$ denotes the event sequence that corresponds to the 
true branch, $ELSE(e_b)$ denotes the event sequence that corresponds to 
the false branch, the branch event $e_b$ is  $if(x_b<y_b)$, the event $e_m$ 
in $IF(e_b)$ is in the form of $x_m=y_m+z_m$ and the 
event $e_n$ in $ELSE(e_b)$ is in the form of $x_n=y_n+z_n$.

{\bf [XZ: missing piece]}

@@

$$
\begin{array}{l}
O'(e_b)<O'(e_{i_1})<O'(e_{i_2})<\dots <O'(e_{i_m})<O'(e_{end})\\
\wedge\\
O'(e_b)<O'(e_{j_1})<O'(e_{j_2})<\dots <O'(e_{j_m})<O'(e_{end})
\end{array}
$$

>>>>

In particular, we have the following criteria in determining if an 
unexecuted branch should be included in the constraint system. These are 
essentially the bound of our path relaxation.
(1) If there is a loop inside the unexecuted branch, the loop bound needs to
be resolved to a constant from the trace (otherwise we do not know how many
times we need to unroll). (2) The unexecuted branch
is ignored if it contains recursive function(s). (3) Every base reference 
variable in the unexecuted branch needs to be fully resolved to a concrete object,
as required by our earlier analysis (Section~\ref{sec:relax1}). This can be 
done by backward traversal of the trace:
if the base reference depends on some shared read in the unexplored 
branch, of which the value is uncertain, it cannot be 
resolved to an object; if the base reference depends on multiple definitions 
from different paths, it cannot be resolved to an object.
Otherwise, it can be resolved.
(4) The indices of array read accesses in the unexecuted branch 
need to be resolved to constant values from the trace. This is to
prevent reading an array element that was not defined. 
(5) The unexecuted branch is ignored if it uses collections. 
(6) For method calls in the unexecuted branch, we require the  caller object can be fully resolved from
the trace.

%We start with a branch event $e_b$ in the trace, without loss of generality, 
%suppose its true branch is taken, we need to explore the false branch 
%(thread-locally with the same thread as $e_b$) and add constraints to model it.
% As the static analysis has some known challenges, we make several 
%simplifying assumptions: 
%(1) the unexplored branch should not contain 
%loops or recursions because the loop bound may depend on some shared variable 
%which is statically unknown; (2) every base reference variable should be 
%fully resolved to a concrete object, as required by our analysis 
%(Section~\ref{sec:relax1}). We apply dataflow analysis (reaching definitions) 
%to judge: If the base reference depends on some shared read in the explored 
%branch, of which the value is statically unknown, it cannot be statically 
%resolved to an object; If the base reference depends on multiple definitions 
%from different paths, it cannot be statically resolved to an object. 
%Otherwise, it can be resolved. (3) We assume the unexplored branch does not 
%contain array access because the array accesses often use the index 
%expressions that are statically unknown. Similarly, as the collections 
%often use array for implementation, we assume they are also absent. (4) We 
%inline the method calls (for two levels) in the unexplored branch, similar 
%to the base object, we assume the caller object can be fully resolved 
%statically.  If any of the above assumptions is violated, we  avoid exploring 
%the unexplored branches and adopt the purely dynamic solution in Section~\ref{sec:relax1}. 
Next, we will explain how to encode an unexecuted branch in details, 
using the example in Figure~\ref{fig:running} for demonstration.
Given a branch event $e_b$, we assume one of its branches is unexecuted and needs to be (statically) extended in
the following discussion.

@@

{\bf SSA Form\ }  Similar to in Section~\ref{sec:relax1}, we need the SSA 
form of the instruction in each event. To achieve this, we apply SSA 
transformation to the code of unexecuted branch and combine its SSA form 
with the existing SSA form of the corresponding executed branch to produce the final 
SSA form. The SSA transformation is standard, but special considerations
are needed.
%issues need to be considered.
\begin{itemize}
\item {\bf Local primitive variables\ }  
For local reads inside the extended branch whose definitions are
%Some local variable is defined 
before the branch event $e_b$, we reuse their SSA forms which 
%of which the SSA form 
are already in the trace $\tau$.
%, we reuse the SSA form.  
In addition, 
%the SSA transformation 
we encode conditional statements inside the extended branch using
Phi node, e.g., $y=Phi(y_1, C_1, y_2, C_2, \dots y_i, C_i)$, 
meaning that $y$ is equal to $y_i$ in the condition $C_i$.
Phi nodes are also used to integrate the definitions in the extended
branch with the executed branch.

@@

{\bf Intra-thread Value Constraint \ } Intra-thread value constraint is the 
core constraint that realizes the exploration of different branches. It 
asserts the value constraints imposed by the {\sf assign/heapr/heapw} events 
take effect only when the guarding condition is satisfied. 
Suppose $IF(e_b)$ denotes the event sequence that corresponds to the 
true branch, $ELSE(e_b)$ denotes the event sequence that corresponds to 
the false branch, the branch event $e_b$ is  $if(x_b<y_b)$, the event $e_m$ 
in $IF(e_b)$ is in the form of $x_m=y_m+z_m$ and the 
event $e_n$ in $ELSE(e_b)$ is in the form of $x_n=y_n+z_n$. Then, we have,
$$
\begin{array}{l}
\bigwedge\limits_{e_m \in IF(e_b)} (\theta'(x_b) < \theta'(y_b) \rightarrow  \theta'(x_m)=\theta'(y_m)+\theta'(z_m)) \\ 
\quad \quad  \bigwedge  \\
\bigwedge\limits_{e_n \in ELSE(e_b)} (\neg(\theta'(x_b)  < \theta'(y_b)) \rightarrow  \theta'(x_n)=\theta'(y_n)+\theta'(z_n)) \\ 
\end{array}
$$

@@

$$
\begin{array}{l}
O'(e_b)<O'(e_{i_1})<O'(e_{i_2})<\dots <O'(e_{i_m})<O'(e_{end})\\
\wedge\\
O'(e_b)<O'(e_{j_1})<O'(e_{j_2})<\dots <O'(e_{j_n})<O'(e_{end})
\end{array}
$$
